Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-12 Manuscript/proof_outline.tex
remove noncomputable

\todo[inline]{omit all noncomputable? Otherwise we need to explain what it means}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def
    projectiveLimitWithWeakestHypotheses
    [∀ i, PseudoEMetricSpace (α i)]
    [∀ i, BorelSpace (α i)]
    [∀ i, SecondCountableTopology (α i)]
    [∀ i, CompleteSpace (α i)] [∀ i, Nonempty (α i)]
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    [∀ i, IsFiniteMeasure (P i)]
    (hP : IsProjectiveMeasureFamily P) :
    Measure (∀ i, α i)
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def diff₀ (hC : SetSemiring C)
    (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
    [DecidableEq (Set α)] : Finset (Set α) :=
    (hC.exists_disjoint_finset_diff_eq hs I hI).choose \ {∅}
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def indexedDiff₀ (hC : SetSemiring C)
    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
    (n : Fin J.card) : Finset (Set α) :=
    hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
    (finsetLT_subset' J hJ n)
\end{minted}

@@

  noncomputable def extendContent
    (hC : SetSemiring C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (m_empty : m ∅ hC.empty_mem = 0)
    (m_add :
      ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
      (h_mem : ⋃₀ ↑I ∈ C),
      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop)) :
    AddContent C 
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def kolmogorovFun
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    (s : Set (∀ i, α i))
    (hs : s ∈ cylinders α) : ℝ≥0∞ :=
    P (cylinders.finset hs) (cylinders.set hs)

  noncomputable def kolContent
    (hP : IsProjectiveMeasureFamily P) :
    AddContent (cylinders α) :=
    extendContent setSemiringCylinders
    (kolmogorovFun P) (kolmogorovFun_empty hP)
    (kolmogorovFun_additive hP)
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def ofAddSubaddCaratheodory
    (hC : SetSemiring C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (m_empty : m ∅ hC.empty_mem = 0)
    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
      (h_mem : ⋃₀ ↑I ∈ C),
      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
      @Measure α (inducedOuterMeasure m
        hC.empty_mem m_empty).caratheodory
\end{minted}
\todo{is this the right lemma? Do we have a similar one about contents?}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
noncomputable def Measure.ofAddContent
  [mα : MeasurableSpace α] (hC : SetSemiring C)
  (hC_gen : MeasurableSpace.generateFrom C = mα)
  (m : AddContent C)
  (m_sigma_subadd : ...) :
  Measure α
\end{minted}
with \leanline{m_sigma_subadd} as above.

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
noncomputable def
  projectiveLimitWithWeakestHypotheses
  (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
  (hP : IsProjectiveMeasureFamily P) :
  Measure (∀ i, α i) :=
    Measure.ofAddContent setSemiringCylinders
    generateFrom_cylinders (kolContent hP)
    (kolContent_countably_subadditive_of_innerRegular hP
      fun J => innerRegular_isCompact_isClosed_
      measurableSet_of_complete_countable (P J))    
\end{minted}
Let us look closer at the formalized proof:
\leanline{kolContent_countably_subadditive_of_innerRegular} has the
same hypotheses as
\leanline{kolContent_sigma_additive_of_innerRegular}, except
\leanline{h_disj}. So, the lemma in the last brackets shows
$\sigma$-sub-additivity of the content $P$ from
Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders} gives the \leanline{MeasurableSpace} on which we define the
\leanline{addContent}.

>>>>

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def projectiveLimitWithWeakestHypotheses
  [∀ i, PseudoEMetricSpace (α i)]
  [∀ i, BorelSpace (α i)]
  [∀ i, SecondCountableTopology (α i)]
  [∀ i, CompleteSpace (α i)] [∀ i, Nonempty (α i)]
  (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
  [∀ i, IsFiniteMeasure (P i)]
  (hP : IsProjectiveMeasureFamily P) :
  Measure (∀ i, α i)
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def diff₀ (hC : SetSemiring C)
    (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
    [DecidableEq (Set α)] : Finset (Set α) :=
  (hC.exists_disjoint_finset_diff_eq hs I hI).choose \ {∅}
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def indexedDiff₀ (hC : SetSemiring C)
    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
    (n : Fin J.card) : Finset (Set α) :=
  hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
    (finsetLT_subset' J hJ n)
\end{minted}

@@

def extendContent
    (hC : SetSemiring C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (m_empty : m ∅ hC.empty_mem = 0)
    (m_add :
      ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
      (h_mem : ⋃₀ ↑I ∈ C),
      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop)) :
    AddContent C 
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def kolmogorovFun
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    (s : Set (∀ i, α i))
    (hs : s ∈ cylinders α) : ℝ≥0∞ :=
  P (cylinders.finset hs) (cylinders.set hs)

def kolContent
    (hP : IsProjectiveMeasureFamily P) :
    AddContent (cylinders α) :=
  extendContent setSemiringCylinders
    (kolmogorovFun P) (kolmogorovFun_empty hP)
    (kolmogorovFun_additive hP)
\end{minted}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def ofAddSubaddCaratheodory
    (hC : SetSemiring C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (m_empty : m ∅ hC.empty_mem = 0)
    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
      (h_mem : ⋃₀ ↑I ∈ C),
      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
    @Measure α (inducedOuterMeasure m
      hC.empty_mem m_empty).caratheodory
\end{minted}
\todo{is this the right lemma? Do we have a similar one about contents?}

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def Measure.ofAddContent
    [mα : MeasurableSpace α] (hC : SetSemiring C)
    (hC_gen : MeasurableSpace.generateFrom C = mα)
    (m : AddContent C)
    (m_sigma_subadd : ...) :
    Measure α
\end{minted}
with \leanline{m_sigma_subadd} as above.

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
def projectiveLimitWithWeakestHypotheses
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    (hP : IsProjectiveMeasureFamily P) :
    Measure (∀ i, α i) :=
  Measure.ofAddContent setSemiringCylinders
    generateFrom_cylinders (kolContent hP)
    (kolContent_countably_subadditive_of_innerRegular hP
      fun J => innerRegular_isCompact_isClosed_
      measurableSet_of_complete_countable (P J))    
\end{minted}
Let us look closer at the formalized proof:
\leanline{kolContent_countably_subadditive_of_innerRegular} has the
same hypotheses as
\leanline{kolContent_sigma_additive_of_innerRegular}, except
\leanline{h_disj}. So, the lemma in the last brackets shows
$\sigma$-sub-additivity of the content $P$ from
Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders} gives the \leanline{MeasurableSpace} on which we define the
\leanline{addContent}.
